---
title: "Wisconsin Gerrymandering"
sidebar: false
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
library(tidyverse)
library(sf)
library(fec16)
library(leaflet)
library(mdsr)
library(mapproj)
library(maps)
library(ggspatial)
library(prettymapr)
library(readr)
WI_Population_Sheet1 <- read_csv("~/WI Population - Sheet1.csv")
```

```{r, include=FALSE}
print(results_house, width = Inf)
results_house |>
  group_by(state, district_id) |>
  summarize(N = n())|>
  print(n=Inf)
```

```{r, include=FALSE, echo=FALSE}
conflicted::conflicts_prefer(base::sum)


district_elections <- results_house |>
  mutate(district = parse_number(district_id)) |>
  group_by(state, district) |>
  summarize(
    N = n(), 
    total_votes = sum(general_votes, na.rm = TRUE),
    d_votes = sum(ifelse(party == "DEM", general_votes, 0), na.rm = TRUE),
    r_votes = sum(ifelse(party == "REP", general_votes, 0), na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    other_votes = total_votes - d_votes - r_votes,
    r_prop = r_votes / total_votes,  
    winner = ifelse(r_votes > d_votes, "Republican", "Democrat")
  )|>
  filter(state=="WI")
wi_results <- district_elections |>
  filter(state == "WI")
wi_results |>                  
  select(-state)
```

```{r, message=FALSE, include=FALSE, echo=FALSE}
src <- "http://cdmaps.polisci.ucla.edu/shp/districts113.zip"
lcl_zip <- fs::path(tempdir(), "districts113.zip")
download.file(src, destfile = lcl_zip)
lcl_districts <- fs::path(tempdir(), "districts113")
unzip(lcl_zip, exdir = lcl_districts)
dsn_districts <- fs::path(lcl_districts, "districtShapes")



st_layers(dsn_districts)
```

```{r, message=FALSE, include=FALSE, echo=FALSE}
districts <- st_read(dsn_districts, layer = "districts113") |>
  mutate(DISTRICT = parse_number(as.character(DISTRICT)))
head(districts, width = Inf)
class(districts)

wi_shp <- districts |>
  filter(STATENAME == "Wisconsin")
wi_shp |>
  st_geometry() |>
  plot(col = gray.colors(nrow(wi_shp)))

wi_merged <- wi_shp |>
  st_transform(4326) |>
  inner_join(wi_results, by = c("DISTRICT" = "district"))

wi <- ggplot(data = wi_merged, aes(fill = winner)) +
  annotation_map_tile(zoom = 6, type = "osm", progress = "none") + 
  geom_sf(alpha = 0.5) +
  scale_fill_manual("Winner", values = c("blue", "red")) + 
  geom_sf_label(aes(label = DISTRICT), fill = "white") + 
  theme_void()
```

```{r, message=FALSE, include=FALSE, echo=FALSE}
wi +
  aes(fill = r_prop) + 
  geom_sf_label(aes(label = DISTRICT), fill = "white")+
  scale_fill_distiller(
    "Proportion\nRepublican", 
    palette = "RdBu", 
    limits = c(0.2, 0.8))
```

```{r, message=FALSE, echo=FALSE}

pal <- colorNumeric(palette = "RdBu", domain = c(0, 1))

leaflet_wi <- leaflet(wi_merged) |>
  addTiles() |>
  addPolygons(
    weight = 1, fillOpacity = 0.8, 
    color = ~pal(1 - r_prop),
    popup = ~paste("District", DISTRICT, "</br>", "Proportion Republican Votes", round(r_prop, 4)))|>
  addPolygons(
    data = wi_merged,
    weight = 0.7, fillOpacity = 0,
    color = "darkgrey",
    popup = ~paste("District", DISTRICT, "</br>", "Proportion Republican Votes", round(r_prop, 4)))|>
  setView(lng = -90, lat = 45, zoom = 6)|>
  addControl(html = '<h2 style="text-align: center;">2016 WI Congressional Districts</h2>',
    position = "topright")

leaflet_wi
```

### 2016 WI Congressional Districts

This map shows the congressional Districts in WI as of 2016. The colors from blue to red represent how respectively Democratic or Republican that district voted in the 2016 elections for Congress.

One interesting aspect is the shapes of the districts. We see that some districts, like Districts 3 & 4 are very oddly shaped. District 3 seems to reach up into areas like Stevens Point and Eau Claire, while District 4 is by far the smallest, and covers little more than Milwaukee. This could bring up accusations of gerrymandering, especially seeing as the two weirdest shaped districts are also 2 of the 3 that elected Democratic Representatives. If, for example, District 4 was a bit bigger, it might draw some of the republican votes away from their respective districts and make those districts more competitive and at risk of turning blue. The size of District 7 is also notable. It is likely that there are some more liberal areas in NW Wisconsin around Superior, however their votes would be cancelled out by the large span of conservative areas in northcentral WI.

We also see that the darkness of the color corresponds to the proportion of votes. For example, districts 3 & 4 have no Republican votes because they were uncontested competitions. The majority of the Republican winning counties had around 60-65% of the vote.

WI Governor Evers recently signed a bill passing a new Legislative Districts map to combat some of this gerrymandering, however thus far the Congressional Districts have remained the same.
